---
title: "Join Lab"
author: "Adam Nur"
date: "DATA 2401"
output: html_document
---


# Joins and `dplyr` practice

Load the `"nycflights13"` package.  You'll also need to load `dplyr` or `tidyverse`, if you didn't above.

```{r}
library(nycflights13)
library(dplyr)

```


Create a dataframe of the average arrival delays for each _destination_

```{r}
avg_arr_delay <- flights %>%
  group_by(dest) %>%
  summarise(avg_arr_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(avg_arr_delay))
print(avg_arr_delay)

```

Then use `left_join()` to join on the "airports" dataframe, which has the airport information.

```{r}
avg_arr_delay_with_info <- avg_arr_delay %>%
  left_join(airports, by = c("dest" = "faa"))
print(avg_arr_delay_with_info)

```


Note here: the column names don't match anymore! 

Which airport had the largest average arrival delay?

```{r}
worst_airport <- avg_arr_delay_with_info %>%
  slice_max(avg_arr_delay, n = 1)

print(worst_airport)

```



What is the average departure delay for each airline? To answer this with something more than just the FAA code of the airline name, you need to create a dataframe of the average departure delay for each _airline_, then use `left_join()` to join on the "airlines" dataframe


```{r}
avg_dep_delay <- flights %>%
  group_by(carrier) %>%
  summarise(avg_dep_delay = mean(dep_delay, na.rm = TRUE)) %>%
  arrange(desc(avg_dep_delay))



avg_dep_delay_with_names <- avg_dep_delay %>%
  left_join(airlines, by = "carrier")

print(avg_dep_delay_with_names)

```

Which airline had the smallest average departure delay? 



```{r}
best_airline <- avg_dep_delay_with_names %>%
  slice_min(avg_dep_delay, n = 1)

print(best_airline)


```





Examples from R4DS:    

```{r}

top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)
top_dest

```

Now you want to find each flight that went to one of those destinations. You could, of course, filter. But! Now you can semi-join: 

```{r}
worst_destinations <- avg_arr_delay %>%
  slice_max(avg_arr_delay, n = 5)

flights_to_worst_destinations <- flights %>%
  semi_join(worst_destinations, by = "dest")

```


What does anti_join(flights, airports, by = c("dest" = "faa")) tell you? What does anti_join(airports, flights, by = c("faa" = "dest")) tell you?
Perform the two joins below and consider what the results are. 
```{r}
unmatched_airports <- airports %>%
  anti_join(flights, by = c("faa" = "dest"))

print(unmatched_airports)

```